---
title: "EDA_analysis2"
format: html
editor: visual
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(screenshot.force = TRUE)
```

```{r}
# Install packages
install.packages("tidyverse", dependencies = TRUE)
install.packages("data.table")
install.packages("dplyr")
install.packages("signal")
install.packages("plotly")
install.packages("pracma")
install.packages("ggpubr")
install.packages("zoo")
install.packages("webshot")

# Download libraries
library(tidyverse)
library(readr)
library(ggplot2)
library(data.table)
library(dplyr)
library(signal)
library(plotly)
library(pracma)
library(ggpubr)
library(zoo)
library(webshot)
```

## Load data sets

```{r}
setwd("~/Desktop/R_docs")

ECG_EDA01 <- fread("ECG_EDA01.csv")
ECG_EDA02 <- fread("ECG_EDA02.csv")
ECG_EDA05 <- fread("ECG_EDA05.csv")
ECG_EDA06 <- fread("ECG_EDA06.csv")
ECG_EDA09 <- fread("ECG_EDA09.csv")
ECG_EDA10 <- fread("ECG_EDA10.csv")
```

## Isolate EDA data per participant

```{r}
# Remove EDA measure
EDA01 <- ECG_EDA01 %>% dplyr::filter(Measure != "ECG")
EDA02 <- ECG_EDA02 %>% dplyr::filter(Measure != "ECG")
EDA05 <- ECG_EDA05 %>% dplyr::filter(Measure != "ECG")
EDA06 <- ECG_EDA06 %>% dplyr::filter(Measure != "ECG")
EDA09 <- ECG_EDA09 %>% dplyr::filter(Measure != "ECG")
EDA10 <- ECG_EDA10 %>% dplyr::filter(Measure != "ECG")
```

## Add Index

```{r}
EDA01 <- EDA01 %>%
  mutate(Index = row_number()) %>% 
  select(Index, everything())

EDA02 <- EDA02 %>%
  mutate(Index = row_number()) %>%
  select(Index, everything())

EDA05 <- EDA05 %>%
  mutate(Index = row_number()) %>%
  select(Index, everything())

EDA06 <- EDA06 %>%
  mutate(Index = row_number()) %>%
  select(Index, everything())

EDA09 <- EDA09 %>%
  mutate(Index = row_number()) %>%
  select(Index, everything())

EDA10 <- EDA10 %>%
  mutate(Index = row_number()) %>%
  select(Index, everything())
```

## Pre-processing Steps

### Smoothing EDA (moving average) signal to reduce noise

#### SCL baseline tracking (long-term trends)

```{r}
# Choose best window_sec size
# Try 3 different windows
EDA01 <- EDA01 %>%
  mutate(
    ma_1s = rollmean(Value, k = 200, fill = NA, align = "right"),
    ma_2s = rollmean(Value, k = 400, fill = NA, align = "right"),
    ma_5s = rollmean(Value, k = 1000, fill = NA, align = "right")
  )

# Plot to compare
plot(EDA01$Timestamp, EDA01$Value, type = "l", col = "gray", main = "Smoothed EDA Signals")
lines(EDA01$Timestamp, EDA01$ma_1s, col = "blue")
lines(EDA01$Timestamp, EDA01$ma_2s, col = "green")
lines(EDA01$Timestamp, EDA01$ma_5s, col = "red")
legend("topright", legend = c("Raw", "1s", "2s", "5s"), col = c("gray", "blue", "green", "red"), lty = 1)

EDA01 <- EDA01 %>%
  select(-ma_1s, -ma_2s, -ma_5s)
```

```{r}
# Setting up the parameters
sampling_rate <- 200  # samples per second 
window_sec <- 5       # smoothing window in seconds
window_samples <- sampling_rate * window_sec  # number of samples in window

# Apply moving average smoothing to raw EDA 'Value'

EDA01 <- EDA01 %>%
  mutate(
    ma_SCL = rollmean(Value, k = window_samples, fill = NA, align = "right")
  )

EDA02 <- EDA02 %>%
  mutate(
    ma_SCL = rollmean(Value, k = window_samples, fill = NA, align = "right")
  )

EDA05 <- EDA05 %>%
  mutate(
    ma_SCL = rollmean(Value, k = window_samples, fill = NA, align = "right")
  )

EDA06 <- EDA06 %>%
  mutate(
    ma_SCL = rollmean(Value, k = window_samples, fill = NA, align = "right")
  )

EDA09 <- EDA09 %>%
  mutate(
    ma_SCL = rollmean(Value, k = window_samples, fill = NA, align = "right")
  )

EDA10 <- EDA10 %>%
  mutate(
    ma_SCL = rollmean(Value, k = window_samples, fill = NA, align = "right")
  )
```

#### SCR baseline tracking (spikes)

```{r}
# Setting up the parameters
sampling_rate <- 200  # Hz
window_sec <- 0.5     # 0.5-second smoothing window
window_samples <- as.integer(sampling_rate * window_sec)  # 100 samples

# Apply light smoothing
EDA01 <- EDA01 %>%
  mutate(
    ma_SCR = rollmean(Value, k = window_samples, fill = NA, align = "right")
  )

EDA02 <- EDA02 %>%
  mutate(
    ma_SCR = rollmean(Value, k = window_samples, fill = NA, align = "right")
  )

EDA05 <- EDA05 %>%
  mutate(
    ma_SCR = rollmean(Value, k = window_samples, fill = NA, align = "right")
  )

EDA06 <- EDA06 %>%
  mutate(
    ma_SCR = rollmean(Value, k = window_samples, fill = NA, align = "right")
  )

EDA09 <- EDA09 %>%
  mutate(
    ma_SCR = rollmean(Value, k = window_samples, fill = NA, align = "right")
  )

EDA10 <- EDA10 %>%
  mutate(
    ma_SCR = rollmean(Value, k = window_samples, fill = NA, align = "right")
  )
```

## Detect R-peaks with findpeaks()

### Create the function detect_eda_peaks

DELETE AFTER: 🔹 Support from literature:

Many studies using **SD-based peak detection** or **z-scoring** set thresholds in the **0.5–1.0 range**. The 0.7 value performs well when: Using **detrended** and **smoothed** EDA; Aiming to detect **acute stress** without overflagging; Sampling rate is **high (e.g., 200 Hz)**, which can otherwise inflate noise –\> Filters moderate-to-large SCRs; balances sensitivity/specificity

🔹 Why 5 seconds? Prevents **multiple peaks** from being counted during a single SCR event by enforcing a **refractory period** between peaks; SCRs have a **rise time** of \~1–3 sec and can take up to **5–10 seconds** to fully resolve; At 200 Hz, very short gaps (e.g., 1–2 sec = 200–400 samples) may double-count **sustained SCRs** or **artifacts**; 5 seconds (1000 samples) is a **conservative** but **safe** choice for clean separation.

This function processes raw EDA data by smoothing the tonic component (SCL), detrending the signal, applying light smoothing for phasic responses (SCR), and detecting SCR peaks based on a standard deviation threshold. It returns a tibble with processed signal components and peak indicators.

```{r}
process_eda_signal <- function(eda_signal, timestamps, 
                               fs = 200,                      # Hz
                               scl_window_sec = 5,            # 5 sec smoothing (SCL)
                               scr_window_sec = 0.5,          # 0.5 sec smoothing (SCR)
                               scr_sd_multiplier = 0.7,       # SD threshold
                               min_scr_gap_sec = 5) {         # Gap between SCRs
  
  # 1. Smooth for tonic SCL (moving average over 5 sec)
  scl_window <- scl_window_sec * fs                           # 1000 samples
  scl <- rollmean(eda_signal, k = scl_window, fill = NA, align = "right")

  # 2. SCR detection: detrend using SCL, then lightly smooth detrended signal
  detrended <- eda_signal - scl
  scr_window <- scr_window_sec * fs                           # 100 samples
  scr_smoothed <- rollmean(detrended, k = scr_window, fill = NA, align = "right")

  # 3. Apply thresholding to detect SCR peaks
  threshold <- sd(scr_smoothed, na.rm = TRUE) * scr_sd_multiplier
  min_gap_samples <- min_scr_gap_sec * fs
  
  peaks <- rep(FALSE, length(eda_signal))
  last_peak <- -Inf

  for (i in seq_len(length(scr_smoothed))) {
    if (!is.na(scr_smoothed[i]) &&
        scr_smoothed[i] > threshold &&
        (i - last_peak) > min_gap_samples) {
      peaks[i] <- TRUE
      last_peak <- i
    }
  }

  # 4. Return tibble
  tibble(
    Time = timestamps,
    Raw_EDA = eda_signal,
    SCL = scl,
    Detrended = detrended,
    SCR_Smoothed = scr_smoothed,
    SCR_Peak = peaks
  )
}
```

### Apply Peak Detection to the Participants

#### Participant 1

```{r}
# Step 1: Smooth signal for SCR detection (0.5-second light smoothing)
sampling_rate <- 200
scr_smoothing_window <- as.integer(sampling_rate * 0.5)  # 100 samples

EDA01 <- EDA01 %>%
  mutate(
    ma_SCR = rollmean(Value, k = scr_smoothing_window, fill = NA, align = "right")
  )

# Step 2: Apply process_eda_signal() with new parameters
processed <- process_eda_signal(
  eda_signal = EDA01$ma_SCR,
  timestamps = EDA01$Timestamp,
  fs = 200,
  scl_window_sec = 5,        # SCL smoothing window
  scr_window_sec = 10,       # SCR detection window
  scr_sd_multiplier = 0.8,   # SCR threshold
  min_scr_gap_sec = 5        # Min gap between SCRs
)

# Step 3: Add SCR peak info back to EDA01
EDA01 <- EDA01 %>%
  select(-any_of("IsPeak")) %>%  # safely remove IsPeak only if it exists
  left_join(
    processed %>% select(Time, SCR_Peak) %>% rename(IsPeak = SCR_Peak),
    by = c("Timestamp" = "Time")
  ) %>%
  mutate(IsPeak = ifelse(is.na(IsPeak), 0, as.integer(IsPeak)))


# Step 4: Create dataset of only peaks
EDA01_peaks <- EDA01 %>% filter(IsPeak == 1)

# Step 5: Check timestamp range
timestamp_range <- range(EDA01$Timestamp, na.rm = TRUE)
cat("Min time:", timestamp_range[1], "\n")
cat("Max time:", timestamp_range[2], "\n")

# Step 6: Plot
plot(EDA01$Timestamp, EDA01$ma_SCR, type = "l", main = "EDA Signal with Detected Peaks", xlab = "Time (s)", ylab = "Smoothed EDA")
points(EDA01_peaks$Timestamp, EDA01_peaks$ma_SCR, col = "red", pch = 20)

```

#### Plot SCL + SCR Peaks

Plotting SCL as a smooth line provides the reference level. Overlaying SCR peaks as points on that line shows when and where the fast responses happen relative to the baseline. If you plotted only SCRs (like just spike markers), you’d lose the overall context of the signal.

```{r}
zoom_start <- 650  
zoom_end <- 850   

# Create zoomed-in subset with peaks
zoomed_data <- EDA01 %>% filter(Timestamp >= zoom_start & Timestamp <= zoom_end)

# Extract peak points within zoom window
zoomed_peaks <- zoomed_data %>% filter(IsPeak == 1)

# Plot smoothed tonic EDA (ma_SCL) with peaks
plot(zoomed_data$Timestamp, zoomed_data$ma_SCL,
     type = "l", col = "blue", xlab = "Time (s)", ylab = "EDA (Smoothed SCL)",
     main = paste("P1 Smoothed EDA Signal with Peaks (Zoom:", zoom_start, "to", zoom_end, "s)"))

points(zoomed_peaks$Timestamp, zoomed_peaks$ma_SCL,
       col = "red", pch = 20)

```

#### Participant 02

```{r}
# Step 1: Smooth signal for SCR detection (0.5-second light smoothing)
sampling_rate <- 200
scr_smoothing_window <- as.integer(sampling_rate * 0.5)  # 100 samples

EDA02 <- EDA02 %>%
  mutate(
    ma_SCR = rollmean(Value, k = scr_smoothing_window, fill = NA, align = "right")
  )

# Step 2: Apply process_eda_signal() with new parameters
processed <- process_eda_signal(
  eda_signal = EDA02$ma_SCR,
  timestamps = EDA02$Timestamp,
  fs = 200,
  scl_window_sec = 5,        # SCL smoothing window
  scr_window_sec = 10,       # SCR detection window
  scr_sd_multiplier = 0.8,   # SCR threshold
  min_scr_gap_sec = 5        # Min gap between SCRs
)

# Step 3: Add SCR peak info back to EDA01
EDA02 <- EDA02 %>%
  select(-any_of("IsPeak")) %>%  # safely remove IsPeak only if it exists
  left_join(
    processed %>% select(Time, SCR_Peak) %>% rename(IsPeak = SCR_Peak),
    by = c("Timestamp" = "Time")
  ) %>%
  mutate(IsPeak = ifelse(is.na(IsPeak), 0, as.integer(IsPeak)))

# Step 4: Create dataset of only peaks
EDA02_peaks <- EDA02 %>% filter(IsPeak == 1)

# Step 5: Check timestamp range
timestamp_range <- range(EDA02$Timestamp, na.rm = TRUE)
cat("Min time:", timestamp_range[1], "\n")
cat("Max time:", timestamp_range[2], "\n")

# Step 6: Plot
plot(EDA02$Timestamp, EDA02$ma_SCR, type = "l", main = "EDA Signal with Detected Peaks", xlab = "Time (s)", ylab = "Smoothed EDA")
points(EDA02_peaks$Timestamp, EDA02_peaks$ma_SCR, col = "red", pch = 20)

```

#### Plot SCL + SCR Peaks

```{r}
zoom_start <- 750  
zoom_end <- 1000    

# Create zoomed-in subset with peaks
zoomed_data <- EDA02 %>% filter(Timestamp >= zoom_start & Timestamp <= zoom_end)

# Extract peak points within zoom window
zoomed_peaks <- zoomed_data %>% filter(IsPeak == 1)

# Plot smoothed tonic EDA (ma_SCL) with peaks
plot(zoomed_data$Timestamp, zoomed_data$ma_SCL,
     type = "l", col = "blue", xlab = "Time (s)", ylab = "EDA (Smoothed SCL)",
     main = paste("Smoothed EDA Signal with Peaks (Zoom:", zoom_start, "to", zoom_end, "s)"))

points(zoomed_peaks$Timestamp, zoomed_peaks$ma_SCL,
       col = "red", pch = 20)

```

### Participant 05

```{r}
# Step 1: Smooth signal for SCR detection (0.5-second light smoothing)
sampling_rate <- 200
scr_smoothing_window <- as.integer(sampling_rate * 0.5)  # 100 samples

EDA05 <- EDA05 %>%
  mutate(
    ma_SCR = rollmean(Value, k = scr_smoothing_window, fill = NA, align = "right")
  )

# Step 2: Apply process_eda_signal() with new parameters
processed <- process_eda_signal(
  eda_signal = EDA05$ma_SCR,
  timestamps = EDA05$Timestamp,
  fs = 200,
  scl_window_sec = 5,        # SCL smoothing window
  scr_window_sec = 10,       # SCR detection window
  scr_sd_multiplier = 0.8,   # SCR threshold
  min_scr_gap_sec = 5        # Min gap between SCRs
)

# Step 3: Add SCR peak info back to EDA01
EDA05 <- EDA05 %>%
  select(-any_of("IsPeak")) %>%  # safely remove IsPeak only if it exists
  left_join(
    processed %>% select(Time, SCR_Peak) %>% rename(IsPeak = SCR_Peak),
    by = c("Timestamp" = "Time")
  ) %>%
  mutate(IsPeak = ifelse(is.na(IsPeak), 0, as.integer(IsPeak)))

# Step 4: Create dataset of only peaks
EDA05_peaks <- EDA05 %>% filter(IsPeak == 1)

# Step 5: Check timestamp range
timestamp_range <- range(EDA05$Timestamp, na.rm = TRUE)
cat("Min time:", timestamp_range[1], "\n")
cat("Max time:", timestamp_range[2], "\n")

# Step 6: Plot
plot(EDA05$Timestamp, EDA05$ma_SCR, type = "l", main = "EDA Signal with Detected Peaks", xlab = "Time (s)", ylab = "Smoothed EDA")
points(EDA05_peaks$Timestamp, EDA05_peaks$ma_SCR, col = "red", pch = 20)

```

#### Plot SCL + SCR Peaks

```{r}
# Plot the data
zoom_start <- 750  
zoom_end <- 1000    

# Create zoomed-in subset with peaks
zoomed_data <- EDA05 %>% filter(Timestamp >= zoom_start & Timestamp <= zoom_end)

# Extract peak points within zoom window
zoomed_peaks <- zoomed_data %>% filter(IsPeak == 1)

# Plot smoothed tonic EDA (ma_SCL) with peaks
plot(zoomed_data$Timestamp, zoomed_data$ma_SCL,
     type = "l", col = "blue", xlab = "Time (s)", ylab = "EDA (Smoothed SCL)",
     main = paste("Smoothed EDA Signal with Peaks (Zoom:", zoom_start, "to", zoom_end, "s)"))

points(zoomed_peaks$Timestamp, zoomed_peaks$ma_SCL,
       col = "red", pch = 20)
```

### Participant 06

```{r}
# Step 1: Smooth signal for SCR detection (0.5-second light smoothing)
sampling_rate <- 200
scr_smoothing_window <- as.integer(sampling_rate * 0.5)  # 100 samples

EDA06 <- EDA06 %>%
  mutate(
    ma_SCR = rollmean(Value, k = scr_smoothing_window, fill = NA, align = "right")
  )

# Step 2: Apply process_eda_signal() with new parameters
processed <- process_eda_signal(
  eda_signal = EDA06$ma_SCR,
  timestamps = EDA06$Timestamp,
  fs = 200,
  scl_window_sec = 5,        # SCL smoothing window
  scr_window_sec = 10,       # SCR detection window
  scr_sd_multiplier = 0.7,   # SCR threshold
  min_scr_gap_sec = 5        # Min gap between SCRs
)

# Step 3: Add SCR peak info back to EDA01
EDA06 <- EDA06 %>%
  select(-any_of("IsPeak")) %>%  # safely remove IsPeak only if it exists
  left_join(
    processed %>% select(Time, SCR_Peak) %>% rename(IsPeak = SCR_Peak),
    by = c("Timestamp" = "Time")
  ) %>%
  mutate(IsPeak = ifelse(is.na(IsPeak), 0, as.integer(IsPeak)))

# Step 4: Create dataset of only peaks
EDA06_peaks <- EDA06 %>% filter(IsPeak == 1)

# Step 5: Check timestamp range
timestamp_range <- range(EDA06$Timestamp, na.rm = TRUE)
cat("Min time:", timestamp_range[1], "\n")
cat("Max time:", timestamp_range[2], "\n")

# Step 6: Plot
plot(EDA06$Timestamp, EDA06$ma_SCR, type = "l", main = "EDA Signal with Detected Peaks", xlab = "Time (s)", ylab = "Smoothed EDA")
points(EDA06_peaks$Timestamp, EDA06_peaks$ma_SCR, col = "red", pch = 20)

```

#### Plot SCL + SCR Peaks

```{r}
# Plot the data
zoom_start <- 750  
zoom_end <- 1000    

# Create zoomed-in subset with peaks
zoomed_data <- EDA06 %>% filter(Timestamp >= zoom_start & Timestamp <= zoom_end)

# Extract peak points within zoom window
zoomed_peaks <- zoomed_data %>% filter(IsPeak == 1)

# Plot smoothed tonic EDA (ma_SCL) with peaks
plot(zoomed_data$Timestamp, zoomed_data$ma_SCL,
     type = "l", col = "blue", xlab = "Time (s)", ylab = "EDA (Smoothed SCL)",
     main = paste("Smoothed EDA Signal with Peaks (Zoom:", zoom_start, "to", zoom_end, "s)"))

points(zoomed_peaks$Timestamp, zoomed_peaks$ma_SCL,
       col = "red", pch = 20)
```

### Participant 09

```{r}
# Step 1: Smooth signal for SCR detection (0.5-second light smoothing)
sampling_rate <- 200
scr_smoothing_window <- as.integer(sampling_rate * 0.5)  # 100 samples

EDA09 <- EDA09 %>%
  mutate(
    ma_SCR = rollmean(Value, k = scr_smoothing_window, fill = NA, align = "right")
  )

# Step 2: Apply process_eda_signal() with new parameters
processed <- process_eda_signal(
  eda_signal = EDA09$ma_SCR,
  timestamps = EDA09$Timestamp,
  fs = 200,
  scl_window_sec = 5,        # SCL smoothing window
  scr_window_sec = 10,       # SCR detection window
  scr_sd_multiplier = 0.7,   # SCR threshold
  min_scr_gap_sec = 5        # Min gap between SCRs
)

# Step 3: Add SCR peak info back to EDA01
EDA09 <- EDA09 %>%
  select(-any_of("IsPeak")) %>%  # safely remove IsPeak only if it exists
  left_join(
    processed %>% select(Time, SCR_Peak) %>% rename(IsPeak = SCR_Peak),
    by = c("Timestamp" = "Time")
  ) %>%
  mutate(IsPeak = ifelse(is.na(IsPeak), 0, as.integer(IsPeak)))

# Step 4: Create dataset of only peaks
EDA09_peaks <- EDA09 %>% filter(IsPeak == 1)

# Step 5: Check timestamp range
timestamp_range <- range(EDA09$Timestamp, na.rm = TRUE)
cat("Min time:", timestamp_range[1], "\n")
cat("Max time:", timestamp_range[2], "\n")

# Step 6: Plot
plot(EDA09$Timestamp, EDA09$ma_SCR, type = "l", main = "EDA Signal with Detected Peaks", xlab = "Time (s)", ylab = "Smoothed EDA")
points(EDA09_peaks$Timestamp, EDA09_peaks$ma_SCR, col = "red", pch = 20)
```

#### Plot SCL + SCR Peaks

```{r}
# Plot the data
zoom_start <- 750  
zoom_end <- 1000    

# Create zoomed-in subset with peaks
zoomed_data <- EDA09 %>% filter(Timestamp >= zoom_start & Timestamp <= zoom_end)

# Extract peak points within zoom window
zoomed_peaks <- zoomed_data %>% filter(IsPeak == 1)

# Plot smoothed tonic EDA (ma_SCL) with peaks
plot(zoomed_data$Timestamp, zoomed_data$ma_SCL,
     type = "l", col = "blue", xlab = "Time (s)", ylab = "EDA (Smoothed SCL)",
     main = paste("Smoothed EDA Signal with Peaks (Zoom:", zoom_start, "to", zoom_end, "s)"))

points(zoomed_peaks$Timestamp, zoomed_peaks$ma_SCL,
       col = "red", pch = 20)
```

### Participant 10

```{r}
# Step 1: Smooth signal for SCR detection (0.5-second light smoothing)
sampling_rate <- 200
scr_smoothing_window <- as.integer(sampling_rate * 0.5)  # 100 samples

EDA10 <- EDA10 %>%
  mutate(
    ma_SCR = rollmean(Value, k = scr_smoothing_window, fill = NA, align = "right")
  )

# Step 2: Apply process_eda_signal() with new parameters
processed <- process_eda_signal(
  eda_signal = EDA10$ma_SCR,
  timestamps = EDA10$Timestamp,
  fs = 200,
  scl_window_sec = 5,        # SCL smoothing window
  scr_window_sec = 10,       # SCR detection window
  scr_sd_multiplier = 0.7,   # SCR threshold
  min_scr_gap_sec = 5        # Min gap between SCRs
)

# Step 3: Add SCR peak info back to EDA01
EDA10 <- EDA10 %>%
  select(-any_of("IsPeak")) %>%  # safely remove IsPeak only if it exists
  left_join(
    processed %>% select(Time, SCR_Peak) %>% rename(IsPeak = SCR_Peak),
    by = c("Timestamp" = "Time")
  ) %>%
  mutate(IsPeak = ifelse(is.na(IsPeak), 0, as.integer(IsPeak)))

# Step 4: Create dataset of only peaks
EDA10_peaks <- EDA10 %>% filter(IsPeak == 1)

# Step 5: Check timestamp range
timestamp_range <- range(EDA10$Timestamp, na.rm = TRUE)
cat("Min time:", timestamp_range[1], "\n")
cat("Max time:", timestamp_range[2], "\n")

# Step 6: Plot
plot(EDA10$Timestamp, EDA10$ma_SCR, type = "l", main = "EDA Signal with Detected Peaks", xlab = "Time (s)", ylab = "Smoothed EDA")
points(EDA10_peaks$Timestamp, EDA10_peaks$ma_SCR, col = "red", pch = 20)
```

#### Plot SCL + SCR Peaks

```{r}
# Plot the data
zoom_start <- 750  
zoom_end <- 1000    

# Create zoomed-in subset with peaks
zoomed_data <- EDA10 %>% filter(Timestamp >= zoom_start & Timestamp <= zoom_end)

# Extract peak points within zoom window
zoomed_peaks <- zoomed_data %>% filter(IsPeak == 1)

# Plot smoothed tonic EDA (ma_SCL) with peaks
plot(zoomed_data$Timestamp, zoomed_data$ma_SCL,
     type = "l", col = "blue", xlab = "Time (s)", ylab = "EDA (Smoothed SCL)",
     main = paste("Smoothed EDA Signal with Peaks (Zoom:", zoom_start, "to", zoom_end, "s)"))

points(zoomed_peaks$Timestamp, zoomed_peaks$ma_SCL,
       col = "red", pch = 20)
```

## Differentiation of Points of Interest in EDA for both SCR and SCL

### Using Moving Average Baseline + Percent Change

The two codes calculate percent change in EDA using moving averages. The **SCR version** uses a 20-second window to capture fast, short-lived arousal spikes typical of phasic responses. The **SCL version** uses a 60-second window to reflect slower, tonic shifts. Both help identify significant changes from recent baseline levels.

```{r}
# Participant 1
EDA01 <- EDA01 %>%
  mutate(
    SCR_ma_baseline = zoo::rollmean(ma_SCR, k = 200 * 20, fill = NA, align = "right"),  # 20-sec baseline
    SCR_ma_change_pct = 100 * (ma_SCR - SCR_ma_baseline) / SCR_ma_baseline
  )

EDA01 <- EDA01 %>%
  mutate(
    SCL_ma_baseline = zoo::rollmean(ma_SCL, k = 200 * 60, fill = NA, align = "right"),  # 60-sec average
    SCL_ma_change_pct = 100 * (ma_SCL - SCL_ma_baseline) / SCL_ma_baseline
  )

# Participant 2
EDA02 <- EDA02 %>%
  mutate(
    SCR_ma_baseline = zoo::rollmean(ma_SCR, k = 200 * 20, fill = NA, align = "right"),  # 20-sec baseline
    SCR_ma_change_pct = 100 * (ma_SCR - SCR_ma_baseline) / SCR_ma_baseline
  )

EDA02 <- EDA02 %>%
  mutate(
    SCL_ma_baseline = zoo::rollmean(ma_SCL, k = 200 * 60, fill = NA, align = "right"),  # 60-sec average
    SCL_ma_change_pct = 100 * (ma_SCL - SCL_ma_baseline) / SCL_ma_baseline
  )

# Participant 5
EDA05 <- EDA05 %>%
 mutate(
    SCR_ma_baseline = zoo::rollmean(ma_SCR, k = 200 * 20, fill = NA, align = "right"),  # 20-sec baseline
    SCR_ma_change_pct = 100 * (ma_SCR - SCR_ma_baseline) / SCR_ma_baseline
  )

EDA05 <- EDA05 %>%
  mutate(
    SCL_ma_baseline = zoo::rollmean(ma_SCL, k = 200 * 60, fill = NA, align = "right"),  # 60-sec average
    SCL_ma_change_pct = 100 * (ma_SCL - SCL_ma_baseline) / SCL_ma_baseline
  )

# Participant 6
EDA06 <- EDA06 %>%
  mutate(
    SCR_ma_baseline = zoo::rollmean(ma_SCR, k = 200 * 20, fill = NA, align = "right"),  # 20-sec baseline
    SCR_ma_change_pct = 100 * (ma_SCR - SCR_ma_baseline) / SCR_ma_baseline
  )

EDA06 <- EDA06 %>%
  mutate(
    SCL_ma_baseline = zoo::rollmean(ma_SCL, k = 200 * 60, fill = NA, align = "right"),  # 60-sec average
    SCL_ma_change_pct = 100 * (ma_SCL - SCL_ma_baseline) / SCL_ma_baseline
  )

# Participant 9
EDA09 <- EDA09 %>%
  mutate(
    SCR_ma_baseline = zoo::rollmean(ma_SCR, k = 200 * 20, fill = NA, align = "right"),  # 20-sec baseline
    SCR_ma_change_pct = 100 * (ma_SCR - SCR_ma_baseline) / SCR_ma_baseline
  )

EDA09 <- EDA09 %>%
  mutate(
    SCL_ma_baseline = zoo::rollmean(ma_SCL, k = 200 * 60, fill = NA, align = "right"),  # 60-sec average
    SCL_ma_change_pct = 100 * (ma_SCL - SCL_ma_baseline) / SCL_ma_baseline
  )

# Participant 10
EDA10 <- EDA10 %>%
  mutate(
    SCR_ma_baseline = zoo::rollmean(ma_SCR, k = 200 * 20, fill = NA, align = "right"),  # 20-sec baseline
    SCR_ma_change_pct = 100 * (ma_SCR - SCR_ma_baseline) / SCR_ma_baseline
  )

EDA10 <- EDA10 %>%
  mutate(
    SCL_ma_baseline = zoo::rollmean(ma_SCL, k = 200 * 60, fill = NA, align = "right"),  # 60-sec average
    SCL_ma_change_pct = 100 * (ma_SCL - SCL_ma_baseline) / SCL_ma_baseline
  )
```

#### Create the Binary Values for Moving Average Baseline USING Z-SCORES

This code creates binary stress indicators based on how much the moving average percent change deviates from the participant’s typical EDA pattern. First, it converts the percent change values for both SCR (short-term arousal) and SCL (long-term arousal) into z-scores, which represent how many standard deviations each value is from the mean. Then, it flags stress (value = 1) when a z-score exceeds +1 —meaning the signal is at least one standard deviation higher than usual, suggesting a significant physiological shift.

```{r}
# Participant 1
EDA01 <- EDA01 %>%
  mutate(
    SCR_ma_change_z = (SCR_ma_change_pct - mean(SCR_ma_change_pct, na.rm = TRUE)) / sd(SCR_ma_change_pct, na.rm = TRUE),
    SCR_ma_binary = ifelse(SCR_ma_change_z >= 1, 1, 0)  # 1 SD threshold for stress
  )

EDA01 <- EDA01 %>%
  mutate(
    SCL_ma_change_z = (SCL_ma_change_pct - mean(SCL_ma_change_pct, na.rm = TRUE)) / sd(SCL_ma_change_pct, na.rm = TRUE),
    SCL_ma_binary = ifelse(SCL_ma_change_z >= 1, 1, 0)  # 1 SD threshold for stress
  )

# Participant 2
EDA02 <- EDA02 %>%
  mutate(
    SCR_ma_change_z = (SCR_ma_change_pct - mean(SCR_ma_change_pct, na.rm = TRUE)) / sd(SCR_ma_change_pct, na.rm = TRUE),
    SCR_ma_binary = ifelse(SCR_ma_change_z >= 1, 1, 0)  # 1 SD threshold for stress
  )

EDA02 <- EDA02 %>%
  mutate(
    SCL_ma_change_z = (SCL_ma_change_pct - mean(SCL_ma_change_pct, na.rm = TRUE)) / sd(SCL_ma_change_pct, na.rm = TRUE),
    SCL_ma_binary = ifelse(SCL_ma_change_z >= 1, 1, 0)  # 1 SD threshold for stress
  )

# Participant 5
EDA05 <- EDA05 %>%
  mutate(
    SCR_ma_change_z = (SCR_ma_change_pct - mean(SCR_ma_change_pct, na.rm = TRUE)) / sd(SCR_ma_change_pct, na.rm = TRUE),
    SCR_ma_binary = ifelse(SCR_ma_change_z >= 1, 1, 0)  # 1 SD threshold for stress
  )

EDA05 <- EDA05 %>%
  mutate(
    SCL_ma_change_z = (SCL_ma_change_pct - mean(SCL_ma_change_pct, na.rm = TRUE)) / sd(SCL_ma_change_pct, na.rm = TRUE),
    SCL_ma_binary = ifelse(SCL_ma_change_z >= 1, 1, 0)  # 1 SD threshold for stress
  )

# Participant 6
EDA06 <- EDA06 %>%
  mutate(
    SCR_ma_change_z = (SCR_ma_change_pct - mean(SCR_ma_change_pct, na.rm = TRUE)) / sd(SCR_ma_change_pct, na.rm = TRUE),
    SCR_ma_binary = ifelse(SCR_ma_change_z >= 1, 1, 0)  # 1 SD threshold for stress
  )

EDA06 <- EDA06 %>%
  mutate(
    SCL_ma_change_z = (SCL_ma_change_pct - mean(SCL_ma_change_pct, na.rm = TRUE)) / sd(SCL_ma_change_pct, na.rm = TRUE),
    SCL_ma_binary = ifelse(SCL_ma_change_z >= 1, 1, 0)  # 1 SD threshold for stress
  )

# Participant 9
EDA09 <- EDA09 %>%
  mutate(
    SCR_ma_change_z = (SCR_ma_change_pct - mean(SCR_ma_change_pct, na.rm = TRUE)) / sd(SCR_ma_change_pct, na.rm = TRUE),
    SCR_ma_binary = ifelse(SCR_ma_change_z >= 1, 1, 0)  # 1 SD threshold for stress
  )

EDA09 <- EDA09 %>%
  mutate(
    SCL_ma_change_z = (SCL_ma_change_pct - mean(SCL_ma_change_pct, na.rm = TRUE)) / sd(SCL_ma_change_pct, na.rm = TRUE),
    SCL_ma_binary = ifelse(SCL_ma_change_z >= 1, 1, 0)  # 1 SD threshold for stress
  )

# Participant 10
EDA10 <- EDA10 %>%
  mutate(
    SCR_ma_change_z = (SCR_ma_change_pct - mean(SCR_ma_change_pct, na.rm = TRUE)) / sd(SCR_ma_change_pct, na.rm = TRUE),
    SCR_ma_binary = ifelse(SCR_ma_change_z >= 1, 1, 0)  # 1 SD threshold for stress
  )

EDA10 <- EDA10 %>%
  mutate(
    SCL_ma_change_z = (SCL_ma_change_pct - mean(SCL_ma_change_pct, na.rm = TRUE)) / sd(SCL_ma_change_pct, na.rm = TRUE),
    SCL_ma_binary = ifelse(SCL_ma_change_z >= 1, 1, 0)  # 1 SD threshold for stress
  )

plot(EDA01$Timestamp, EDA01$SCR_ma_change_pct, type = "l", main = "EDA01 SCR Exceeding 1 SD")
points(EDA01$Timestamp[EDA01$SCR_ma_binary == 1], 
       EDA01$SCR_ma_change_pct[EDA01$SCR_ma_binary == 1], 
       col = "red", pch = 20)

plot(EDA01$Timestamp, EDA01$SCL_ma_change_pct, type = "l", main = "EDA01 SCL Exceeding 1 SD")
points(EDA01$Timestamp[EDA01$SCL_ma_binary == 1], 
       EDA01$SCL_ma_change_pct[EDA01$SCL_ma_binary == 1], 
       col = "red", pch = 20)

```

```{r}
# Function to print counts and percentages for a binary column in a dataframe
print_binary_percentages <- function(df, col_name) {
  cat(paste0("Counts and percentages for ", col_name, ":\n"))
  counts <- table(df[[col_name]])
  total <- sum(counts)
  percentages <- round(100 * counts / total, 2)
  result <- data.frame(Count = counts, Percentage = percentages)
  print(result)
  cat("\n")
}

# Apply to your EDA01 dataset
print_binary_percentages(EDA01, "SCR_ma_binary")
print_binary_percentages(EDA01, "SCL_ma_binary")
print_binary_percentages(EDA02, "SCR_ma_binary")
print_binary_percentages(EDA02, "SCL_ma_binary")
print_binary_percentages(EDA05, "SCR_ma_binary")
print_binary_percentages(EDA05, "SCL_ma_binary")
print_binary_percentages(EDA06, "SCR_ma_binary")
print_binary_percentages(EDA06, "SCL_ma_binary")
print_binary_percentages(EDA09, "SCR_ma_binary")
print_binary_percentages(EDA09, "SCL_ma_binary")
print_binary_percentages(EDA10, "SCR_ma_binary")
print_binary_percentages(EDA10, "SCL_ma_binary")
```

### Rising Trend Detector Using Moving Average Baseline (SCR_onset) USING Z-SCORES

This function detects acute stress onset by looking for a rising trend in the z-scored percent change of short-term SCR activity. It flags the start of a stress episode when the SCR signal increases compared to the previous value and exceeds a defined threshold (default = 0.7). Once stress is detected, it continues labeling until the signal starts to drop. This method works well for SCR because it captures fast, phasic shifts. SCL, by contrast, isn’t well-suited for onset detection since its slower, tonic changes lack sharp or time-locked increases.

```{r}
function_stress_onset <- function(df, pct_col = "SCR_ma_change_z", 
                                 output_col = "SCR_onset", threshold = 0.7) {
  stress_label <- integer(nrow(df))
  in_stress <- FALSE
  
  pct_values <- df[[pct_col]]
  
  for (i in 2:length(pct_values)) {
    if (!in_stress) {
      if (!is.na(pct_values[i]) && !is.na(pct_values[i - 1]) &&
          (pct_values[i] > pct_values[i - 1]) &&
          (pct_values[i] >= threshold)) {
        in_stress <- TRUE
        stress_label[i] <- 1
      }
    } else {
      stress_label[i] <- 1
      if (!is.na(pct_values[i]) && !is.na(pct_values[i - 1]) &&
          (pct_values[i] < pct_values[i - 1])) {
        in_stress <- FALSE
      }
    }
  }
  
  df[[output_col]] <- stress_label
  return(df)
}

```

#### Apply the Function to All Participants

```{r}
EDA01 <- function_stress_onset(EDA01, pct_col = "SCR_ma_change_z", output_col = "SCR_onset", threshold = 0.7)
EDA02 <- function_stress_onset(EDA02, pct_col = "SCR_ma_change_z", output_col = "SCR_onset", threshold = 0.7)
EDA05 <- function_stress_onset(EDA05, pct_col = "SCR_ma_change_z", output_col = "SCR_onset", threshold = 0.7)
EDA06 <- function_stress_onset(EDA06, pct_col = "SCR_ma_change_z", output_col = "SCR_onset", threshold = 0.7)
EDA09 <- function_stress_onset(EDA09, pct_col = "SCR_ma_change_z", output_col = "SCR_onset", threshold = 0.7)
EDA10 <- function_stress_onset(EDA10, pct_col = "SCR_ma_change_z", output_col = "SCR_onset", threshold = 0.7)
```

```{r}
library(ggplot2)

ggplot(EDA01, aes(x = Timestamp)) +
  geom_line(aes(y = SCR_ma_change_z), color = "steelblue", size = 0.6) +
  geom_point(aes(y = ifelse(SCR_onset == 1, SCR_ma_change_z, NA)),
           color = "red", size = 2, alpha = 0.7) +
labs(title = "P1 Stress Onset Detection Based on SCR Moving Average % Change (≥ 1 SD)",
       x = "Time (s)",
       y = "SCR (z-scored moving avg change)") +
  theme_minimal()

```

#### Percentage of 1s and 0s

```{r}
percent_ones_zeros <- function(df, col) {
  total <- nrow(df)
  ones <- sum(df[[col]] == 1, na.rm = TRUE)
  zeros <- sum(df[[col]] == 0, na.rm = TRUE)
  nas <- sum(is.na(df[[col]]))
  percent_ones <- ones / total * 100
  percent_zeros <- zeros / total * 100
  percent_nas <- nas / total * 100
  return(data.frame(Percent_1s = percent_ones, Percent_0s = percent_zeros, Percent_NA = percent_nas))
}


# Apply to each dataset
percent_ones_zeros(EDA01, "SCR_onset") # 1=0.97 0=99
percent_ones_zeros(EDA02, "SCR_onset") # 1=1.2 0=98.8
percent_ones_zeros(EDA05, "SCR_onset") # 1=0.48 0=99.5
percent_ones_zeros(EDA06, "SCR_onset") # 1=0.82 0=99
percent_ones_zeros(EDA09, "SCR_onset") # 1=0.72 0=99.3 
percent_ones_zeros(EDA10, "SCR_onset") # 1=0 0=100

```

#### Create a new data set with stress onset data

Marks the **start of a rising phase** and labels `1`s until the trend reverses. The labeled `1`s form **blocks** in your data → these are **stress events** (start to end of rising response).

```{r}
EDA01_stress_onset_SCR <- EDA01 %>%
  filter(SCR_onset == 1)

EDA02_stress_onset_SCR <- EDA02 %>%
  filter(SCR_onset == 1)

EDA05_stress_onset_SCR <- EDA05 %>%
  filter(SCR_onset == 1)

EDA06_stress_onset_SCR <- EDA06 %>%
  filter(SCR_onset == 1)

EDA09_stress_onset_SCR <- EDA09 %>%
  filter(SCR_onset == 1)

EDA10_stress_onset_SCR <- EDA10 %>%
  filter(SCR_onset == 1)
```

### Sustained Stress based on SCL USING Z-SCORES

This function detects sustained stress periods based on slow-changing SCL trends. It uses a rolling mean of the z-scored percent change in SCL (`SCL_ma_change_z`) over a minimum duration (e.g., 2 seconds) to smooth out noise. If the rolling average stays above a defined threshold (default = 0.10), the function flags that period as stress. This approach is well-suited to SCL because it captures prolonged arousal states rather than sharp spikes.

```{r}
function_ma_Sustained_stress_duration <- function(df,
                                               pct_col = "SCL_ma_change_z",
                                               output_col = "SCL_sustained",
                                               sampling_rate = 200,
                                               min_duration_sec = 2,
                                               threshold = 0.10) {
  window_size <- sampling_rate * min_duration_sec
  z_values <- df[[pct_col]]  # use pct_col here
  
  rolling_mean <- zoo::rollmean(z_values, k = window_size, fill = NA, align = "right")
  stress_label <- integer(length(z_values))
  
  in_stress <- FALSE
  
  for (i in seq_along(rolling_mean)) {
    if (!in_stress) {
      if (!is.na(rolling_mean[i]) && rolling_mean[i] >= threshold) {
        in_stress <- TRUE
        stress_label[i] <- 1
      }
    } else {
      if (!is.na(rolling_mean[i]) && rolling_mean[i] >= threshold) {
        stress_label[i] <- 1
      } else {
        in_stress <- FALSE
      }
    }
  }
  
  df[[output_col]] <- stress_label
  return(df)
}
```

#### Apply the Function to All Participants

This code applies a sustained stress detection function to the Participant SCL signal. It looks for periods where the z-scored change in SCL (`SCL_ma_change_z`) stays above a threshold of 1.0 for at least 2 seconds (based on a 200 Hz sampling rate). These periods are flagged as stress in a new column called `SCL_sustained`. The output includes both the count and percentage of time points marked as sustained stress vs. non-stress, helping assess how frequently this criterion is met.

```{r}
# Participant 1
EDA01 <- function_ma_Sustained_stress_duration(
  df = EDA01,
  pct_col = "SCL_ma_change_z",     # column with z-scored SCL changes
  output_col = "SCL_sustained",    # new column to store stress labels
  sampling_rate = 200,             # your data's sampling rate (Hz)
  min_duration_sec = 2,            # minimum duration for sustained stress (seconds)
  threshold = 1.0                 # threshold for z-scored SCL changes
)

# Check on how many stress episodes detected
stress_counts <- table(EDA01$SCL_sustained)
stress_percent <- prop.table(stress_counts) * 100

# Print counts and percentages nicely
cat("Counts of sustained stress labels:\n")
print(stress_counts)
cat("\nPercentages of sustained stress labels:\n")
print(round(stress_percent, 2))

# Participant 2
EDA02 <- function_ma_Sustained_stress_duration(
  df = EDA02,
  pct_col = "SCL_ma_change_z",     # column with z-scored SCL changes
  output_col = "SCL_sustained",    # new column to store stress labels
  sampling_rate = 200,             # your data's sampling rate (Hz)
  min_duration_sec = 2,            # minimum duration for sustained stress (seconds)
  threshold = 1.0                 # threshold for z-scored SCL changes
)

# Check on how many stress episodes detected
stress_counts <- table(EDA02$SCL_sustained)
stress_percent <- prop.table(stress_counts) * 100

# Print counts and percentages nicely
cat("Counts of sustained stress labels:\n")
print(stress_counts)
cat("\nPercentages of sustained stress labels:\n")
print(round(stress_percent, 2))

# Participant 5
EDA05 <- function_ma_Sustained_stress_duration(
  df = EDA05,
  pct_col = "SCL_ma_change_z",     # column with z-scored SCL changes
  output_col = "SCL_sustained",    # new column to store stress labels
  sampling_rate = 200,             # your data's sampling rate (Hz)
  min_duration_sec = 2,            # minimum duration for sustained stress (seconds)
  threshold = 1.0                 # threshold for z-scored SCL changes
)

# Check on how many stress episodes detected
stress_counts <- table(EDA05$SCL_sustained)
stress_percent <- prop.table(stress_counts) * 100

# Print counts and percentages nicely
cat("Counts of sustained stress labels:\n")
print(stress_counts)
cat("\nPercentages of sustained stress labels:\n")
print(round(stress_percent, 2))

# Participant 6
EDA06 <- function_ma_Sustained_stress_duration(
  df = EDA06,
  pct_col = "SCL_ma_change_z",     # column with z-scored SCL changes
  output_col = "SCL_sustained",    # new column to store stress labels
  sampling_rate = 200,             # your data's sampling rate (Hz)
  min_duration_sec = 2,            # minimum duration for sustained stress (seconds)
  threshold = 1.0                 # threshold for z-scored SCL changes
)

# Check on how many stress episodes detected
stress_counts <- table(EDA06$SCL_sustained)
stress_percent <- prop.table(stress_counts) * 100

# Print counts and percentages nicely
cat("Counts of sustained stress labels:\n")
print(stress_counts)
cat("\nPercentages of sustained stress labels:\n")
print(round(stress_percent, 2))

# Participant 9
EDA09 <- function_ma_Sustained_stress_duration(
  df = EDA09,
  pct_col = "SCL_ma_change_z",     # column with z-scored SCL changes
  output_col = "SCL_sustained",    # new column to store stress labels
  sampling_rate = 200,             # your data's sampling rate (Hz)
  min_duration_sec = 2,            # minimum duration for sustained stress (seconds)
  threshold = 1.0                 # threshold for z-scored SCL changes
)

# Check on how many stress episodes detected
stress_counts <- table(EDA09$SCL_sustained)
stress_percent <- prop.table(stress_counts) * 100

# Print counts and percentages nicely
cat("Counts of sustained stress labels:\n")
print(stress_counts)
cat("\nPercentages of sustained stress labels:\n")
print(round(stress_percent, 2))

# Participant 2
EDA10 <- function_ma_Sustained_stress_duration(
  df = EDA10,
  pct_col = "SCL_ma_change_z",     # column with z-scored SCL changes
  output_col = "SCL_sustained",    # new column to store stress labels
  sampling_rate = 200,             # your data's sampling rate (Hz)
  min_duration_sec = 2,            # minimum duration for sustained stress (seconds)
  threshold = 1.0                 # threshold for z-scored SCL changes
)

# Check on how many stress episodes detected
stress_counts <- table(EDA10$SCL_sustained)
stress_percent <- prop.table(stress_counts) * 100

# Print counts and percentages nicely
cat("Counts of sustained stress labels:\n")
print(stress_counts)
cat("\nPercentages of sustained stress labels:\n")
print(round(stress_percent, 2))

```

```{r}
# Example assumes your data is in EDA01
ggplot(EDA01, aes(x = Timestamp)) +
  geom_line(aes(y = SCL_ma_change_z), color = "steelblue", linewidth = 0.6) +
  geom_point(
    data = subset(EDA01, SCL_sustained == 1),
    aes(y = SCL_ma_change_z),
    color = "red", size = 1.5, alpha = 0.8
  ) +
  labs(
    title = "P1 Sustained Stress Detection from SCL Signal",
    x = "Time (s)",
    y = "Z-scored % Change in SCL"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14),
    axis.title = element_text(size = 12)
  )

```

#### Create a new data set with sustained stress data

This dataset, `EDA01_stress_sustained_SCL`, contains **only the time points from Participant 1’s data** where **sustained stress** was detected based on SCL. Specifically, it filters rows where the `SCL_sustained` column equals `1`, meaning the SCL signal showed a **z-scored increase ≥ 1.0** sustained for **at least 2 seconds** (as defined in your stress detection function).

In short, this dataset isolates **moments of prolonged physiological arousal** according to your **SCL-based sustained stress criterion**, making it useful for further inspection, visualization, or comparing with other stress indicators.

```{r}
EDA01_stress_sustained_SCL <- EDA01 %>%
  filter(SCL_sustained == 1)

EDA02_stress_sustained_SCL <- EDA02 %>%
  filter(SCL_sustained == 1)

EDA05_stress_sustained_SCL <- EDA05 %>%
  filter(SCL_sustained == 1)

EDA06_stress_sustained_SCL <- EDA06 %>%
  filter(SCL_sustained == 1)

EDA09_stress_sustained_SCL <- EDA09 %>%
  filter(SCL_sustained == 1)

EDA10_stress_sustained_SCL <- EDA10 %>%
  filter(SCL_sustained == 1)
```

### Using baseline: Identifying the Differentiation % per Participant

This code standardizes the SCR signal relative to a defined baseline period. It calculates the mean and standard deviation of the baseline, then z-scores the full signal to create `SCR_baseline_change_z`. Time points exceeding 1.5 standard deviations are flagged as stress events. The output includes a count of flagged points and a plot showing the z-scored signal with threshold and detected stress periods overlaid.

#### Participant 01

```{r}
# Step 1: Define Baseline Windows (in seconds)
baseline1_start <- 161
baseline1_end   <- 478
baseline2_start <- 1371
baseline2_end   <- 1665

# Step 2: Extract baseline SCR_baseline_change values
EDA01_baseline_z_scores <- EDA01 %>%
  filter((OG_Timestamp >= baseline1_start & OG_Timestamp <= baseline1_end) |
         (OG_Timestamp >= baseline2_start & OG_Timestamp <= baseline2_end)) %>%
  pull(ma_SCR)

# Step 3: Compute baseline mean and standard deviation
eda_baseline_mean <- mean(EDA01_baseline_z_scores, na.rm = TRUE)
eda_baseline_sd   <- sd(EDA01_baseline_z_scores, na.rm = TRUE)

# Step 4: Apply baseline-based z-score to full dataset
EDA01 <- EDA01 %>%
  mutate(SCR_baseline_change_z = (ma_SCR - eda_baseline_mean) / eda_baseline_sd)

# Step 5: Define SD-based threshold (e.g., 1 SD above baseline)
sd_multiplier <- 0.7
eda_threshold_sd <- sd_multiplier  # z-score threshold directly

# Step 6: Add stress flag based on new z-score column
EDA01 <- EDA01 %>%
  mutate(
    EDA_baseline_mean = eda_baseline_mean,
    EDA_baseline_sd = eda_baseline_sd,
    EDA_threshold_SD = eda_threshold_sd,
    SCR_baseline_stress_binary = ifelse(SCR_baseline_change_z >= eda_threshold_sd, 1, 0)
  )

# Step 7: Create filtered dataset of flagged stress events
EDA01_baseline_stress_events <- EDA01 %>%
  filter(SCR_baseline_stress_binary == 1)

# Step 8: Count number of stress flags
flag_counts <- table(EDA01$SCR_baseline_stress_binary)
print(flag_counts)
flag_percentages <- prop.table(flag_counts) * 100
print(round(flag_percentages, 2))

# Step 9: Plot for visual check
plot(EDA01$OG_Timestamp, EDA01$SCR_baseline_change_z, type = "l",
     main = "P1 SCR (Z) with SD-Based Threshold",
     ylab = "SCR_baseline_change_z", xlab = "Time (s)")
abline(h = eda_threshold_sd, col = "red", lty = 2)
points(EDA01$OG_Timestamp[EDA01$SCR_baseline_stress_binary == 1],
       EDA01$SCR_baseline_change_z[EDA01$SCR_baseline_stress_binary == 1],
       col = "red", pch = 20)
```

Checking whether SCR_ma_change_z is well-standardised.

```{r}
# Define your baseline time range
baseline1_start <- 161
baseline1_end   <- 478
baseline2_start <- 1371
baseline2_end   <- 1665

# Filter the SCR_ma_change_z values during the baseline period
baseline_z_scores <- EDA01 %>%
  filter((OG_Timestamp >= baseline1_start & OG_Timestamp <= baseline1_end) |
         (OG_Timestamp >= baseline2_start & OG_Timestamp <= baseline2_end)) %>%
  pull(SCR_baseline_change_z)

# Check mean and standard deviation
mean_z <- mean(baseline_z_scores, na.rm = TRUE)
sd_z   <- sd(baseline_z_scores, na.rm = TRUE)

# Print results
cat("Baseline mean of SCR_ma_change_z:", round(mean_z, 3), "\n")
cat("Baseline SD of SCR_ma_change_z:  ", round(sd_z, 3), "\n")
```

#### Participant 02

```{r}
# Step 1: Define Baseline Windows (in seconds)
baseline1_start <- 161
baseline1_end   <- 478
baseline2_start <- 1371
baseline2_end   <- 1665

# Step 2: Extract baseline SCR_baseline_change values
EDA02_baseline_z_scores <- EDA02 %>%
  filter((OG_Timestamp >= baseline1_start & OG_Timestamp <= baseline1_end) |
         (OG_Timestamp >= baseline2_start & OG_Timestamp <= baseline2_end)) %>%
  pull(ma_SCR)

# Step 3: Compute baseline mean and standard deviation
eda_baseline_mean <- mean(EDA02_baseline_z_scores, na.rm = TRUE)
eda_baseline_sd   <- sd(EDA02_baseline_z_scores, na.rm = TRUE)

# Step 4: Apply baseline-based z-score to full dataset
EDA02 <- EDA02 %>%
  mutate(SCR_baseline_change_z = (ma_SCR - eda_baseline_mean) / eda_baseline_sd)

# Step 5: Define SD-based threshold (e.g., 1.5 SD above baseline)
sd_multiplier <- 0.7
eda_threshold_sd <- sd_multiplier  # z-score threshold directly

# Step 6: Add stress flag based on new z-score column
EDA02 <- EDA02 %>%
  mutate(
    EDA_baseline_mean = eda_baseline_mean,
    EDA_baseline_sd = eda_baseline_sd,
    EDA_threshold_SD = eda_threshold_sd,
    SCR_baseline_stress_binary = ifelse(SCR_baseline_change_z >= eda_threshold_sd, 1, 0)
  )

# Step 7: Create filtered dataset of flagged stress events
EDA02_baseline_stress_events <- EDA02 %>%
  filter(SCR_baseline_stress_binary == 1)

# Step 8: Count number of stress flags
flag_counts <- table(EDA02$SCR_baseline_stress_binary)
print(flag_counts)
flag_percentages <- prop.table(flag_counts) * 100
print(round(flag_percentages, 2))

# Step 9: Plot for visual check
plot(EDA02$OG_Timestamp, EDA02$SCR_baseline_change_z, type = "l",
     main = "SCR (Z) with SD-Based Threshold",
     ylab = "SCR_baseline_change_z", xlab = "Time (s)")
abline(h = eda_threshold_sd, col = "red", lty = 2)
points(EDA02$OG_Timestamp[EDA02$SCR_baseline_stress_binary == 1],
       EDA02$SCR_baseline_change_z[EDA02$SCR_baseline_stress_binary == 1],
       col = "red", pch = 20)
```

#### Participant 05

```{r}
# Step 1: Define Baseline Windows (in seconds)
baseline1_start <- 156
baseline1_end   <- 466
baseline2_start <- 1368
baseline2_end   <- 1657

# Step 2: Extract baseline SCR_baseline_change values
EDA05_baseline_z_scores <- EDA05 %>%
  filter((OG_Timestamp >= baseline1_start & OG_Timestamp <= baseline1_end) |
         (OG_Timestamp >= baseline2_start & OG_Timestamp <= baseline2_end)) %>%
  pull(ma_SCR)

# Step 3: Compute baseline mean and standard deviation
eda_baseline_mean <- mean(EDA05_baseline_z_scores, na.rm = TRUE)
eda_baseline_sd   <- sd(EDA05_baseline_z_scores, na.rm = TRUE)

# Step 4: Apply baseline-based z-score to full dataset
EDA05 <- EDA05 %>%
  mutate(SCR_baseline_change_z = (ma_SCR - eda_baseline_mean) / eda_baseline_sd)

# Step 5: Define SD-based threshold (e.g., 1.5 SD above baseline)
sd_multiplier <- 0.7
eda_threshold_sd <- sd_multiplier  # z-score threshold directly

# Step 6: Add stress flag based on new z-score column
EDA05 <- EDA05 %>%
  mutate(
    EDA_baseline_mean = eda_baseline_mean,
    EDA_baseline_sd = eda_baseline_sd,
    EDA_threshold_SD = eda_threshold_sd,
    SCR_baseline_stress_binary = ifelse(SCR_baseline_change_z >= eda_threshold_sd, 1, 0)
  )

# Step 7: Create filtered dataset of flagged stress events
EDA05_baseline_stress_events <- EDA05 %>%
  filter(SCR_baseline_stress_binary == 1)

# Step 8: Count number of stress flags
flag_counts <- table(EDA05$SCR_baseline_stress_binary)
print(flag_counts)
flag_percentages <- prop.table(flag_counts) * 100
print(round(flag_percentages, 2))

# Step 9: Plot for visual check
plot(EDA05$OG_Timestamp, EDA05$SCR_baseline_change_z, type = "l",
     main = "SCR (Z) with SD-Based Threshold",
     ylab = "SCR_baseline_change_z", xlab = "Time (s)")
abline(h = eda_threshold_sd, col = "red", lty = 2)
points(EDA05$OG_Timestamp[EDA05$SCR_baseline_stress_binary == 1],
       EDA05$SCR_baseline_change_z[EDA05$SCR_baseline_stress_binary == 1],
       col = "red", pch = 20)
```

#### Participant 06

```{r}
# Step 1: Define Baseline Windows (in seconds)
baseline1_start <- 156
baseline1_end   <- 466
baseline2_start <- 1368
baseline2_end   <- 1657

# Step 2: Extract baseline SCR_baseline_change values
EDA06_baseline_z_scores <- EDA06 %>%
  filter((OG_Timestamp >= baseline1_start & OG_Timestamp <= baseline1_end) |
         (OG_Timestamp >= baseline2_start & OG_Timestamp <= baseline2_end)) %>%
  pull(ma_SCR)

# Step 3: Compute baseline mean and standard deviation
eda_baseline_mean <- mean(EDA06_baseline_z_scores, na.rm = TRUE)
eda_baseline_sd   <- sd(EDA06_baseline_z_scores, na.rm = TRUE)

# Step 4: Apply baseline-based z-score to full dataset
EDA06 <- EDA06 %>%
  mutate(SCR_baseline_change_z = (ma_SCR - eda_baseline_mean) / eda_baseline_sd)

# Step 5: Define SD-based threshold
sd_multiplier <- 0.7
eda_threshold_sd <- sd_multiplier  # z-score threshold directly

# Step 6: Add stress flag based on new z-score column
EDA06 <- EDA06 %>%
  mutate(
    EDA_baseline_mean = eda_baseline_mean,
    EDA_baseline_sd = eda_baseline_sd,
    EDA_threshold_SD = eda_threshold_sd,
    SCR_baseline_stress_binary = ifelse(SCR_baseline_change_z >= eda_threshold_sd, 1, 0)
  )

# Step 7: Create filtered dataset of flagged stress events
EDA06_baseline_stress_events <- EDA06 %>%
  filter(SCR_baseline_stress_binary == 1)

# Step 8: Count number of stress flags
flag_counts <- table(EDA06$SCR_baseline_stress_binary)
print(flag_counts)
flag_percentages <- prop.table(flag_counts) * 100
print(round(flag_percentages, 2))

# Step 9: Plot for visual check
plot(EDA06$OG_Timestamp, EDA06$SCR_baseline_change_z, type = "l",
     main = "SCR (Z) with SD-Based Threshold",
     ylab = "SCR_baseline_change_z", xlab = "Time (s)")
abline(h = eda_threshold_sd, col = "red", lty = 2)
points(EDA06$OG_Timestamp[EDA06$SCR_baseline_stress_binary == 1],
       EDA06$SCR_baseline_change_z[EDA06$SCR_baseline_stress_binary == 1],
       col = "red", pch = 20)
```

#### Participant 09

```{r}
# Step 1: Define Baseline Windows (in seconds)
baseline1_start <- 161
baseline1_end   <- 473
baseline2_start <- 1580
baseline2_end   <- 1879

# Step 2: Extract baseline SCR_baseline_change values
EDA09_baseline_z_scores <- EDA09 %>%
  filter((OG_Timestamp >= baseline1_start & OG_Timestamp <= baseline1_end) |
         (OG_Timestamp >= baseline2_start & OG_Timestamp <= baseline2_end)) %>%
  pull(ma_SCR)

# Step 3: Compute baseline mean and standard deviation
eda_baseline_mean <- mean(EDA09_baseline_z_scores, na.rm = TRUE)
eda_baseline_sd   <- sd(EDA09_baseline_z_scores, na.rm = TRUE)

# Step 4: Apply baseline-based z-score to full dataset
EDA09 <- EDA09 %>%
  mutate(SCR_baseline_change_z = (ma_SCR - eda_baseline_mean) / eda_baseline_sd)

# Step 5: Define SD-based threshold
sd_multiplier <- 0.7
eda_threshold_sd <- sd_multiplier  # z-score threshold directly

# Step 6: Add stress flag based on new z-score column
EDA09 <- EDA09 %>%
  mutate(
    EDA_baseline_mean = eda_baseline_mean,
    EDA_baseline_sd = eda_baseline_sd,
    EDA_threshold_SD = eda_threshold_sd,
    SCR_baseline_stress_binary = ifelse(SCR_baseline_change_z >= eda_threshold_sd, 1, 0)
  )

# Step 7: Create filtered dataset of flagged stress events
EDA09_baseline_stress_events <- EDA09 %>%
  filter(SCR_baseline_stress_binary == 1)

# Step 8: Count number of stress flags
flag_counts <- table(EDA09$SCR_baseline_stress_binary)
print(flag_counts)
flag_percentages <- prop.table(flag_counts) * 100
print(round(flag_percentages, 2))

# Step 9: Plot for visual check
plot(EDA09$OG_Timestamp, EDA09$SCR_baseline_change_z, type = "l",
     main = "SCR (Z) with SD-Based Threshold",
     ylab = "SCR_baseline_change_z", xlab = "Time (s)")
abline(h = eda_threshold_sd, col = "red", lty = 2)
points(EDA09$OG_Timestamp[EDA09$SCR_baseline_stress_binary == 1],
       EDA09$SCR_baseline_change_z[EDA09$SCR_baseline_stress_binary == 1],
       col = "red", pch = 20)
```

#### Participant 10

```{r}
# Step 1: Define Baseline Windows (in seconds)
baseline1_start <- 161
baseline1_end   <- 473
baseline2_start <- 1580
baseline2_end   <- 1879

# Step 2: Extract baseline SCR_baseline_change values
EDA10_baseline_z_scores <- EDA10 %>%
  filter((OG_Timestamp >= baseline1_start & OG_Timestamp <= baseline1_end) |
         (OG_Timestamp >= baseline2_start & OG_Timestamp <= baseline2_end)) %>%
  pull(ma_SCR)

# Step 3: Compute baseline mean and standard deviation
eda_baseline_mean <- mean(EDA10_baseline_z_scores, na.rm = TRUE)
eda_baseline_sd   <- sd(EDA10_baseline_z_scores, na.rm = TRUE)

# Step 4: Apply baseline-based z-score to full dataset
EDA10 <- EDA10 %>%
  mutate(SCR_baseline_change_z = (ma_SCR - eda_baseline_mean) / eda_baseline_sd)

# Step 5: Define SD-based threshold
sd_multiplier <- 0.7
eda_threshold_sd <- sd_multiplier  # z-score threshold directly

# Step 6: Add stress flag based on new z-score column
EDA10 <- EDA10 %>%
  mutate(
    EDA_baseline_mean = eda_baseline_mean,
    EDA_baseline_sd = eda_baseline_sd,
    EDA_threshold_SD = eda_threshold_sd,
    SCR_baseline_stress_binary = ifelse(SCR_baseline_change_z >= eda_threshold_sd, 1, 0)
  )

# Step 7: Create filtered dataset of flagged stress events
EDA10_baseline_stress_events <- EDA10 %>%
  filter(SCR_baseline_stress_binary == 1)

# Step 8: Count number of stress flags
flag_counts <- table(EDA10$SCR_baseline_stress_binary)
print(flag_counts)
flag_percentages <- prop.table(flag_counts) * 100
print(round(flag_percentages, 2))

# Step 9: Plot for visual check
plot(EDA10$OG_Timestamp, EDA10$SCR_baseline_change_z, type = "l",
     main = "SCR (Z) with SD-Based Threshold",
     ylab = "SCR_baseline_change_z", xlab = "Time (s)")
abline(h = eda_threshold_sd, col = "red", lty = 2)
points(EDA10$OG_Timestamp[EDA10$SCR_baseline_stress_binary == 1],
       EDA10$SCR_baseline_change_z[EDA10$SCR_baseline_stress_binary == 1],
       col = "red", pch = 20)
```

### Label Sustained Stress Duration using Baseline Percent Change

```{r}
function_SCL_sustained_stress_z <- function(df, 
                                     z_col = "SCL_baseline_change_z", 
                                     sampling_rate = 200, 
                                     min_duration_sec = 60, 
                                     threshold = 0.7) {
  
  window_size <- sampling_rate * min_duration_sec
  z_values <- df[[z_col]]
  stress_label <- integer(length(z_values))
  
  in_stress <- FALSE
  counter <- 0
  
  for (i in seq_along(z_values)) {
    if (!in_stress) {
      if (!is.na(z_values[i]) && z_values[i] >= threshold) {
        counter <- counter + 1
        if (counter >= window_size) {
          stress_label[(i - window_size + 1):i] <- 1
          in_stress <- TRUE
        }
      } else {
        counter <- 0
      }
    } else {
      if (!is.na(z_values[i]) && z_values[i] >= threshold) {
        stress_label[i] <- 1
      } else {
        in_stress <- FALSE
        counter <- 0
      }
    }
  }
  
  df$SCL_baseline_sustained <- stress_label
  return(df)
}

```

#### Apply to Participant 1

```{r}
# Step 1: Define baseline windows and calculate baseline mean + SD
baseline1 <- EDA01 %>% filter(OG_Timestamp >= 161 & OG_Timestamp <= 478)
baseline2 <- EDA01 %>% filter(OG_Timestamp >= 1371 & OG_Timestamp <= 1665)

scl_baseline_values <- as.numeric(c(baseline1$ma_SCL, baseline2$ma_SCL))
scl_baseline_mean <- mean(scl_baseline_values, na.rm = TRUE)
scl_baseline_sd   <- sd(scl_baseline_values, na.rm = TRUE)


# Step 2: Compute z-scores for the entire dataset
EDA01 <- EDA01 %>%
  mutate(SCL_baseline_change_z = (ma_SCL - scl_baseline_mean) / scl_baseline_sd)

# Step 3: Label sustained stress if z-score >= 0.7 for 60 seconds
EDA01 <- function_SCL_sustained_stress_z(
  df = EDA01,
  z_col = "SCL_baseline_change_z",
  sampling_rate = 200,
  min_duration_sec = 60,
  threshold = 0.7
)

# Count how many points were flagged
flag_counts <- table(EDA01$SCL_baseline_sustained)
flag_percentages <- prop.table(flag_counts) * 100
print(round(flag_percentages, 2))

```

```{r}
ggplot(EDA01, aes(x = Timestamp)) +
  geom_line(aes(y = SCL_baseline_change_z), color = "black", linewidth = 0.6) +
  geom_point(
    data = subset(EDA01, SCL_baseline_sustained == 1),
    aes(y = SCL_baseline_change_z),
    color = "red", size = 1.5, alpha = 0.8
  ) +
  labs(
    title = "P1 Sustained Stress Detection (≥60s) from SCL Baseline Z-scores",
    x = "Time (s)",
    y = "Z-scored % Change from Baseline (SCL)"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14, face = "bold"),
    axis.title = element_text(size = 12)
  )
```

#### Apply to Participant 2

```{r}
# Step 1: Define baseline windows and calculate baseline mean + SD
baseline1 <- EDA02 %>% filter(OG_Timestamp >= 161 & OG_Timestamp <= 478)
baseline2 <- EDA02 %>% filter(OG_Timestamp >= 1371 & OG_Timestamp <= 1665)

scl_baseline_values <- as.numeric(c(baseline1$ma_SCL, baseline2$ma_SCL))
scl_baseline_mean <- mean(scl_baseline_values, na.rm = TRUE)
scl_baseline_sd   <- sd(scl_baseline_values, na.rm = TRUE)


# Step 2: Compute z-scores for the entire dataset
EDA02 <- EDA02 %>%
  mutate(SCL_baseline_change_z = (ma_SCL - scl_baseline_mean) / scl_baseline_sd)

# Step 3: Label sustained stress if z-score >= 0.7 for 60 seconds
EDA02 <- function_SCL_sustained_stress_z(
  df = EDA02,
  z_col = "SCL_baseline_change_z",
  sampling_rate = 200,
  min_duration_sec = 60,
  threshold = 0.7
)

# Count how many points were flagged
flag_counts <- table(EDA02$SCL_baseline_sustained)
flag_percentages <- prop.table(flag_counts) * 100
print(round(flag_percentages, 2))

```

#### Apply to Participant 5

```{r}
# Step 1: Calculate baseline mean from defined baseline windows
baseline1 <- EDA05 %>% filter(OG_Timestamp >= 156 & OG_Timestamp <= 466)
baseline2 <- EDA05 %>% filter(OG_Timestamp >= 1368 & OG_Timestamp <= 1657)

scl_baseline_values <- as.numeric(c(baseline1$ma_SCL, baseline2$ma_SCL))
scl_baseline_mean <- mean(scl_baseline_values, na.rm = TRUE)
scl_baseline_sd   <- sd(scl_baseline_values, na.rm = TRUE)


# Step 2: Compute z-scores for the entire dataset
EDA05 <- EDA05 %>%
  mutate(SCL_baseline_change_z = (ma_SCL - scl_baseline_mean) / scl_baseline_sd)

# Step 3: Label sustained stress if z-score >= 0.7 for 60 seconds
EDA05 <- function_SCL_sustained_stress_z(
  df = EDA05,
  z_col = "SCL_baseline_change_z",
  sampling_rate = 200,
  min_duration_sec = 60,
  threshold = 0.7
)

# Count how many points were flagged
flag_counts <- table(EDA05$SCL_baseline_sustained)
flag_percentages <- prop.table(flag_counts) * 100
print(round(flag_percentages, 2))

```

#### Apply to Participant 6

```{r}
# Step 1: Calculate baseline mean from defined baseline windows
baseline1 <- EDA06 %>% filter(OG_Timestamp >= 156 & OG_Timestamp <= 466)
baseline2 <- EDA06 %>% filter(OG_Timestamp >= 1368 & OG_Timestamp <= 1657)

scl_baseline_values <- as.numeric(c(baseline1$ma_SCL, baseline2$ma_SCL))
scl_baseline_mean <- mean(scl_baseline_values, na.rm = TRUE)
scl_baseline_sd   <- sd(scl_baseline_values, na.rm = TRUE)

# Step 2: Compute z-scores for the entire dataset
EDA06 <- EDA06 %>%
  mutate(SCL_baseline_change_z = (ma_SCL - scl_baseline_mean) / scl_baseline_sd)

# Step 3: Label sustained stress if z-score >= 0.7 for 60 seconds
EDA06 <- function_SCL_sustained_stress_z(
  df = EDA06,
  z_col = "SCL_baseline_change_z",
  sampling_rate = 200,
  min_duration_sec = 60,
  threshold = 0.7
)

# Count how many points were flagged
flag_counts <- table(EDA06$SCL_baseline_sustained)
flag_percentages <- prop.table(flag_counts) * 100
print(round(flag_percentages, 2))

```

#### Apply to Participant 9

```{r}
# Step 1: Calculate baseline mean from defined baseline windows
baseline1 <- EDA09 %>% filter(OG_Timestamp >= 161 & OG_Timestamp <= 473)
baseline2 <- EDA09 %>% filter(OG_Timestamp >= 1580 & OG_Timestamp <= 1879)

scl_baseline_values <- as.numeric(c(baseline1$ma_SCL, baseline2$ma_SCL))
scl_baseline_mean <- mean(scl_baseline_values, na.rm = TRUE)
scl_baseline_sd   <- sd(scl_baseline_values, na.rm = TRUE)


# Step 2: Compute z-scores for the entire dataset
EDA09 <- EDA09 %>%
  mutate(SCL_baseline_change_z = (ma_SCL - scl_baseline_mean) / scl_baseline_sd)

# Step 3: Label sustained stress if z-score >= 0.7 for 60 seconds
EDA09 <- function_SCL_sustained_stress_z(
  df = EDA09,
  z_col = "SCL_baseline_change_z",
  sampling_rate = 200,
  min_duration_sec = 60,
  threshold = 0.7
)

# Count how many points were flagged
flag_counts <- table(EDA09$SCL_baseline_sustained)
flag_percentages <- prop.table(flag_counts) * 100
print(round(flag_percentages, 2))
```

#### Apply to Participant 10

```{r}
# Step 1: Calculate baseline mean from defined baseline windows
baseline1 <- EDA10 %>% filter(OG_Timestamp >= 161 & OG_Timestamp <= 473)
baseline2 <- EDA10 %>% filter(OG_Timestamp >= 1580 & OG_Timestamp <= 1879)

scl_baseline_values <- as.numeric(c(baseline1$ma_SCL, baseline2$ma_SCL))
scl_baseline_mean <- mean(scl_baseline_values, na.rm = TRUE)
scl_baseline_sd   <- sd(scl_baseline_values, na.rm = TRUE)

# Step 2: Compute z-scores for the entire dataset
EDA10 <- EDA10 %>%
  mutate(SCL_baseline_change_z = (ma_SCL - scl_baseline_mean) / scl_baseline_sd)

# Step 3: Label sustained stress if z-score >= 0.7 for 60 seconds
EDA10 <- function_SCL_sustained_stress_z(
  df = EDA10,
  z_col = "SCL_baseline_change_z",
  sampling_rate = 200,
  min_duration_sec = 60,
  threshold = 0.7
)

# Count how many points were flagged
flag_counts <- table(EDA10$SCL_baseline_sustained)
flag_percentages <- prop.table(flag_counts) * 100
print(round(flag_percentages, 2))
```

### Filter Sustained Stress Durations

This dataset includes only the time points where sustained stress was detected based on baseline-corrected SCL levels. Specifically, it filters `EDA01` to keep rows where `SCL_baseline_sustained` equals 1, indicating periods where SCL remained elevated (based on z-scored deviation from baseline) for a minimum duration—suggesting a tonic stress response.

```{r}
EDA01_SCL_baseline_sustained <- EDA01 %>% filter(SCL_baseline_sustained == 1)
EDA02_SCL_baseline_sustained <- EDA02 %>% filter(SCL_baseline_sustained == 1)
EDA05_SCL_baseline_sustained <- EDA05 %>% filter(SCL_baseline_sustained == 1)
EDA06_SCL_baseline_sustained <- EDA06 %>% filter(SCL_baseline_sustained == 1)
EDA09_SCL_baseline_sustained <- EDA09 %>% filter(SCL_baseline_sustained == 1)
EDA10_SCL_baseline_sustained <- EDA10 %>% filter(SCL_baseline_sustained == 1)
```

## FFT Measures

### Create the function for FFT calculations

This code computes the dominant frequency of the z-scored skin conductance response (SCR) signal using a Fast Fourier Transform (FFT) applied over rolling 10-second windows. It identifies the strongest frequency component in each window and labels periods as stress (`FFT_stress_flag = 1`) if the dominant frequency exceeds 0.5 Hz—a threshold associated with increased sympathetic arousal. This frequency-based method captures brief, phasic stress responses that may not be evident in amplitude-based metrics.

window_sec = 2 step_sec = 0.5

```         
More sensitive to short bursts

Better aligns with phasic SCR events (which typically last ~1–4 seconds)
```

```{r}
apply_fft_rolling <- function(df, signal_col, sampling_rate = 200, 
                              window_sec = 10, step_sec = 0.5) {
  
  window_size <- window_sec * sampling_rate
  step_size <- step_sec * sampling_rate
  n <- nrow(df)
  
  dominant_freqs <- rep(NA, n)
  
  for (start_idx in seq(1, n - window_size, by = step_size)) {
    end_idx <- start_idx + window_size - 1
    segment <- df[[signal_col]][start_idx:end_idx]
    
    if (any(is.na(segment))) {
      # Skip window if NA
      next
    }
    
    # Detrend and apply Hanning window
    segment <- segment - mean(segment)
    segment <- segment * hanning(length(segment))
    
    fft_result <- abs(fft(segment))
    freqs <- seq(0, sampling_rate / 2, length.out = length(fft_result) / 2 + 1)
    power <- fft_result[1:(length(fft_result)/2 + 1)]
    
    dominant_idx <- which.max(power)
    dominant_freqs[start_idx + floor(window_size / 2)] <- freqs[dominant_idx]
    
    # Debug print for first few iterations only
    if (start_idx == 1 || start_idx == step_size * 2) {
      print(paste("Window", start_idx, "dominant freq:", freqs[dominant_idx]))
    }
  }
  
  df$dominant_freq <- dominant_freqs
  return(df)
}

function_participant_eda_fixed_threshold <- function(df, 
                                                    signal_col = "ECG_signal", 
                                                    sampling_rate = 200, 
                                                    window_sec = 10, 
                                                    step_sec = 0.5, 
                                                    fixed_threshold = 0.25) {
  # Apply rolling FFT analysis
  df <- apply_fft_rolling(df, signal_col, sampling_rate, window_sec, step_sec)
  
  # Apply fixed frequency threshold
  cat("Fixed frequency threshold: ", fixed_threshold, " Hz\n")
  
  # Classify stress based on dominant frequency
  df <- df %>%
    mutate(FFT_binary = ifelse(dominant_freq > fixed_threshold, 1, 0))
  
  # Print summary
  cat("FFT-based stress flag counts:\n")
  print(table(df$FFT_binary, useNA = "ifany"))
  
  cat("Percentages:\n")
  print(round(prop.table(table(df$FFT_binary, useNA = "ifany")) * 100, 2))
  
  return(df)
}

```

### Apply the FFT function to the Participants

#### Participant 1

```{r}
# Step 2: Apply the function to EDA01
EDA01 <- function_participant_eda_fixed_threshold(
  EDA01,
  signal_col = "SCR_baseline_change_z",
  sampling_rate = 207,
  window_sec = 7,
  step_sec = 0.25,
  fixed_threshold = 0.25
)

# Define scale factor outside
scale_factor <- 5

# Create scaled dominant frequency column
EDA01$dominant_freq_scaled <- EDA01$dominant_freq * scale_factor

ggplot(EDA01, aes(x = Timestamp)) +
  geom_line(aes(y = SCR_baseline_change_z), color = "black", alpha = 0.5) +
  geom_point(aes(y = dominant_freq_scaled, color = as.factor(FFT_binary)),
             size = 1.5, alpha = 0.7, na.rm = TRUE) +
  geom_vline(data = EDA01[EDA01$FFT_binary == 1, ],
             aes(xintercept = Timestamp), color = "red", alpha = 0.05) +
  scale_color_manual(values = c("0" = "blue", "1" = "red"),
                     labels = c("Relaxed", "Stress")) +
  labs(title = "P1 EDA Signal with FFT-based Stress Labels",
       y = "EDA Z-Score / Scaled Dominant Freq", color = "FFT Binary Label") +
  theme_minimal()

```

#### Participant 2

```{r}
# Step 2: Apply the function to EDA2
EDA02 <- function_participant_eda_fixed_threshold(
  EDA02,
  signal_col = "SCR_baseline_change_z",
  sampling_rate = 207,
  window_sec = 7,
  step_sec = 0.25,
  fixed_threshold = 0.25
)

# Define scale factor outside
scale_factor <- 5

# Create scaled dominant frequency column
EDA02$dominant_freq_scaled <- EDA02$dominant_freq * scale_factor

ggplot(EDA02, aes(x = Timestamp)) +
  geom_line(aes(y = SCR_baseline_change_z), color = "black", alpha = 0.5) +
  geom_point(aes(y = dominant_freq_scaled, color = as.factor(FFT_binary)),
             size = 1.5, alpha = 0.7, na.rm = TRUE) +
  geom_vline(data = EDA01[EDA01$FFT_binary == 1, ],
             aes(xintercept = Timestamp), color = "red", alpha = 0.05) +
  scale_color_manual(values = c("0" = "blue", "1" = "red"),
                     labels = c("Relaxed", "Stress")) +
  labs(title = "P2 EDA Signal with FFT-based Stress Labels",
       y = "EDA Z-Score / Scaled Dominant Freq", color = "FFT Binary Label") +
  theme_minimal()

```

#### Participant 5

```{r}
# Step 2: Apply the function to ECG01
EDA05 <- function_participant_eda_fixed_threshold(
  EDA05,
  signal_col = "SCR_baseline_change_z",
  sampling_rate = 207,
  window_sec = 7,
  step_sec = 0.25,
  fixed_threshold = 0.25
)

# Define scale factor outside
scale_factor <- 5

# Create scaled dominant frequency column
EDA05$dominant_freq_scaled <- EDA05$dominant_freq * scale_factor

ggplot(EDA05, aes(x = Timestamp)) +
  geom_line(aes(y = SCR_baseline_change_z), color = "black", alpha = 0.5) +
  geom_point(aes(y = dominant_freq_scaled, color = as.factor(FFT_binary)),
             size = 1.5, alpha = 0.7, na.rm = TRUE) +
  geom_vline(data = EDA01[EDA01$FFT_binary == 1, ],
             aes(xintercept = Timestamp), color = "red", alpha = 0.05) +
  scale_color_manual(values = c("0" = "blue", "1" = "red"),
                     labels = c("Relaxed", "Stress")) +
  labs(title = "P5 EDA Signal with FFT-based Stress Labels",
       y = "EDA Z-Score / Scaled Dominant Freq", color = "FFT Binary Label") +
  theme_minimal()

```

#### Participant 6

```{r}
# Step 2: Apply the function to ECG01
EDA06 <- function_participant_eda_fixed_threshold(
  EDA06,
  signal_col = "SCR_baseline_change_z",
  sampling_rate = 207,
  window_sec = 7,
  step_sec = 0.25,
  fixed_threshold = 0.25
)

# Define scale factor outside
scale_factor <- 5

# Create scaled dominant frequency column
EDA06$dominant_freq_scaled <- EDA06$dominant_freq * scale_factor

ggplot(EDA06, aes(x = Timestamp)) +
  geom_line(aes(y = SCR_baseline_change_z), color = "black", alpha = 0.5) +
  geom_point(aes(y = dominant_freq_scaled, color = as.factor(FFT_binary)),
             size = 1.5, alpha = 0.7, na.rm = TRUE) +
  geom_vline(data = EDA01[EDA01$FFT_binary == 1, ],
             aes(xintercept = Timestamp), color = "red", alpha = 0.05) +
  scale_color_manual(values = c("0" = "blue", "1" = "red"),
                     labels = c("Relaxed", "Stress")) +
  labs(title = "P6 EDA Signal with FFT-based Stress Labels",
       y = "EDA Z-Score / Scaled Dominant Freq", color = "FFT Binary Label") +
  theme_minimal()

```

#### Participant 9

```{r}
# Step 2: Apply the function to ECG01
EDA09 <- function_participant_eda_fixed_threshold(
  EDA09,
  signal_col = "SCR_baseline_change_z",
  sampling_rate = 207,
  window_sec = 7,
  step_sec = 0.25,
  fixed_threshold = 0.25
)

# Define scale factor outside
scale_factor <- 5

# Create scaled dominant frequency column
EDA09$dominant_freq_scaled <- EDA09$dominant_freq * scale_factor

ggplot(EDA09, aes(x = Timestamp)) +
  geom_line(aes(y = SCR_baseline_change_z), color = "black", alpha = 0.5) +
  geom_point(aes(y = dominant_freq_scaled, color = as.factor(FFT_binary)),
             size = 1.5, alpha = 0.7, na.rm = TRUE) +
  geom_vline(data = EDA01[EDA01$FFT_binary == 1, ],
             aes(xintercept = Timestamp), color = "red", alpha = 0.05) +
  scale_color_manual(values = c("0" = "blue", "1" = "red"),
                     labels = c("Relaxed", "Stress")) +
  labs(title = "EDA Signal with FFT-based Stress Labels",
       y = "EDA Z-Score / Scaled Dominant Freq", color = "FFT Binary Label") +
  theme_minimal()

```

#### Participant 10

```{r}
# Step 2: Apply the function to ECG01
EDA10 <- function_participant_eda_fixed_threshold(
  EDA10,
  signal_col = "SCR_baseline_change_z",
  sampling_rate = 207,
  window_sec = 7,
  step_sec = 0.25,
  fixed_threshold = 0.25
)

# Define scale factor outside
scale_factor <- 5

# Create scaled dominant frequency column
EDA10$dominant_freq_scaled <- EDA10$dominant_freq * scale_factor

ggplot(EDA10, aes(x = Timestamp)) +
  geom_line(aes(y = SCR_baseline_change_z), color = "black", alpha = 0.5) +
  geom_point(aes(y = dominant_freq_scaled, color = as.factor(FFT_binary)),
             size = 1.5, alpha = 0.7, na.rm = TRUE) +
  geom_vline(data = EDA01[EDA01$FFT_binary == 1, ],
             aes(xintercept = Timestamp), color = "red", alpha = 0.05) +
  scale_color_manual(values = c("0" = "blue", "1" = "red"),
                     labels = c("Relaxed", "Stress")) +
  labs(title = "P1 EDA Signal with FFT-based Stress Labels",
       y = "EDA Z-Score / Scaled Dominant Freq", color = "FFT Binary Label") +
  theme_minimal()

```

```{r}
library(ggplot2)
library(patchwork)

# (a) SCR_ma_change_pct with SCR_ma_binary overlay
plot_a <- ggplot(EDA01, aes(x = Timestamp)) +
  geom_line(aes(y = SCR_ma_change_pct), color = "steelblue", linewidth = 0.6) +
  geom_point(
    data = subset(EDA01, SCR_ma_binary == 1),
    aes(y = SCR_ma_change_pct),
    color = "red", size = 1.5, alpha = 0.8
  ) +
  labs(
    title = "(a) EDA01 SCR Exceeding 1 SD",
    x = "Time (s)",
    y = "% Change in SCR"
  ) +
  theme_minimal()

# (b) SCL_ma_change_pct with SCL_ma_binary overlay
plot_b <- ggplot(EDA01, aes(x = Timestamp)) +
  geom_line(aes(y = SCL_ma_change_pct), color = "steelblue", linewidth = 0.6) +
  geom_point(
    data = subset(EDA01, SCL_ma_binary == 1),
    aes(y = SCL_ma_change_pct),
    color = "red", size = 1.5, alpha = 0.8
  ) +
  labs(
    title = "(b) EDA01 SCL Exceeding 1 SD",
    x = "Time (s)",
    y = "% Change in SCL"
  ) +
  theme_minimal()

# (c) SCR moving average z-score with onset
plot_c <- ggplot(EDA01, aes(x = Timestamp)) +
  geom_line(aes(y = SCR_ma_change_z), color = "steelblue", linewidth = 0.6) +
  geom_point(
    data = subset(EDA01, SCR_onset == 1),
    aes(y = SCR_ma_change_z),
    color = "red", size = 2, alpha = 0.7
  ) +
  labs(
    title = "(c) SCR Moving Avg % Change (z) with Stress Onset",
    x = "Time (s)",
    y = "SCR z-score"
  ) +
  theme_minimal()

# (d) SCL moving average z-score with sustained flag
plot_d <- ggplot(EDA01, aes(x = Timestamp)) +
  geom_line(aes(y = SCL_ma_change_z), color = "steelblue", linewidth = 0.6) +
  geom_point(
    data = subset(EDA01, SCL_sustained == 1),
    aes(y = SCL_ma_change_z),
    color = "red", size = 1.5, alpha = 0.8
  ) +
  labs(
    title = "(d) SCL Moving Avg % Change (z) with Sustained Stress",
    x = "Time (s)",
    y = "SCL z-score"
  ) +
  theme_minimal()

# Combine with patchwork
combined_plot <- (plot_a + plot_b) / (plot_c + plot_d)

# Print all 4 panels
combined_plot


# Set up 1 row, 2 columns layout
par(mfrow = c(1, 2), mar = c(4, 4, 4, 2))

```

```{r}
# Save to PNG with custom width and height
png("SCR_facet_P1_P2.png", width = 1800, height = 800, res = 150)

# Set up side-by-side layout
par(mfrow = c(1, 2), mar = c(4, 4, 3, 1))  # margins: bottom, left, top, right

# (a) P1
plot(EDA01$OG_Timestamp, EDA01$SCR_baseline_change_z, type = "l",
     main = "P1 SCR (Z) with SD-Based Threshold",
     ylab = "SCR_baseline_change_z", xlab = "Time (s)")
abline(h = eda_threshold_sd, col = "red", lty = 2)
points(EDA01$OG_Timestamp[EDA01$SCR_baseline_stress_binary == 1],
       EDA01$SCR_baseline_change_z[EDA01$SCR_baseline_stress_binary == 1],
       col = "red", pch = 20)
mtext("(a)", side = 3, line = 1, adj = 0, cex = 1.2)

# (b) P2
plot(EDA02$OG_Timestamp, EDA02$SCR_baseline_change_z, type = "l",
     main = "P2 SCR (Z) with SD-Based Threshold",
     ylab = "SCR_baseline_change_z", xlab = "Time (s)")
abline(h = eda_threshold_sd, col = "red", lty = 2)
points(EDA02$OG_Timestamp[EDA02$SCR_baseline_stress_binary == 1],
       EDA02$SCR_baseline_change_z[EDA02$SCR_baseline_stress_binary == 1],
       col = "red", pch = 20)
mtext("(b)", side = 3, line = 1, adj = 0, cex = 1.2)

dev.off()

```

## Merge participants into an All_EDA

```{r}
# Add ParticipantID to each dataset
EDA01 <- EDA01 %>% mutate(ParticipantID = "01")
EDA02 <- EDA02 %>% mutate(ParticipantID = "02")
EDA05 <- EDA05 %>% mutate(ParticipantID = "05")
EDA06 <- EDA06 %>% mutate(ParticipantID = "06")
EDA09 <- EDA09 %>% mutate(ParticipantID = "09")
EDA10 <- EDA10 %>% mutate(ParticipantID = "10")

# Merge all datasets
All_EDA <- bind_rows(EDA01, EDA02, EDA05, EDA06, EDA09, EDA10)

# Arrange and relocate ParticipantID
All_EDA <- All_EDA_PostFeedback %>%
  arrange(ParticipantID, Timestamp)


All_EDA <- All_EDA %>%
  relocate(ParticipantID, .after = OG_Timestamp)
```

## Save all data sets

### EDA0X

```{r}
write.csv(EDA01, "EDA01.csv", row.names = FALSE)
write.csv(EDA02, "EDA02.csv", row.names = FALSE)
write.csv(EDA05, "EDA05.csv", row.names = FALSE)
write.csv(EDA06, "EDA06.csv", row.names = FALSE)
write.csv(EDA09, "EDA09.csv", row.names = FALSE)
write.csv(EDA10, "EDA10.csv", row.names = FALSE)

write.csv(All_EDA, "All_EDA_PostFeedback.csv", row.names = FALSE)
```

### EDA0X_peaks

```{r}
write.csv(EDA01_peaks, "EDA01_peaks.csv", row.names = FALSE)
write.csv(EDA02_peaks, "EDA02_peaks.csv", row.names = FALSE)
write.csv(EDA05_peaks, "EDA05_peaks.csv", row.names = FALSE)
write.csv(EDA06_peaks, "EDA06_peaks.csv", row.names = FALSE)
write.csv(EDA09_peaks, "EDA09_peaks.csv", row.names = FALSE)
write.csv(EDA10_peaks, "EDA10_peaks.csv", row.names = FALSE)
```

### EDA0X_stress_points

```{r}
write.csv(EDA01_stress_pts, "EDA01_stress_points.csv", row.names = FALSE)
write.csv(EDA02_stress_pts, "EDA02_stress_points.csv", row.names = FALSE)
write.csv(EDA05_stress_pts, "EDA05_stress_points.csv", row.names = FALSE)
write.csv(EDA06_stress_pts, "EDA06_stress_points.csv", row.names = FALSE)
write.csv(EDA09_stress_pts, "EDA09_stress_points.csv", row.names = FALSE)
write.csv(EDA10_stress_pts, "EDA10_stress_points.csv", row.names = FALSE)
```

### EDA0X_stress_events_baseline

```{r}
write.csv(EDA01_stress_events_baseline, "EDA01_stress_events_baseline.csv", row.names = FALSE)
write.csv(EDA02_stress_events_baseline, "EDA02_stress_events_baseline.csv", row.names = FALSE)
write.csv(EDA05_stress_events_baseline, "EDA05_stress_events_baseline.csv", row.names = FALSE)
write.csv(EDA06_stress_events_baseline, "EDA06_stress_events_baseline.csv", row.names = FALSE)
write.csv(EDA09_stress_events_baseline, "EDA09_stress_events_baseline.csv", row.names = FALSE)
write.csv(EDA10_stress_events_baseline, "EDA10_stress_events_baseline.csv", row.names = FALSE)
```

### EDA0X_stress_episodes

```{r}
write.csv(EDA01_stress_episodes, "EDA01_stress_episodes.csv", row.names = FALSE)
write.csv(EDA02_stress_episodes, "EDA02_stress_episodes.csv", row.names = FALSE)
write.csv(EDA05_stress_episodes, "EDA05_stress_episodes.csv", row.names = FALSE)
write.csv(EDA06_stress_episodes, "EDA06_stress_episodes.csv", row.names = FALSE)
write.csv(EDA09_stress_episodes, "EDA09_stress_episodes.csv", row.names = FALSE)
write.csv(EDA10_stress_episodes, "EDA10_stress_episodes.csv", row.names = FALSE)
```
